---
title: "NARW Occurrence GAM Model Testing"
format: html
self-contained: true
editor: visual
fig-width: 12
fig-height: 12
out-width: "100%"
toc: true
knitr:
    opts_chunk:
        echo: true
        collapse: false
        message: false
        warning: false
        comment: "#>"
---

# Packages

```{r}
rm(list = ls())
library(dplyr)
library(lubridate)
# library(scales)

# Visualization
library(ggplot2)
theme_set(theme_light())
library(patchwork)
library(GGally)

# Modeling
library(gamlss)
# library(gamlss.tr)
library(gamlss.add)
library(mgcv)
```


# Load Monthly Data

```{r}
RWocc0 <- read.csv("RWEnv_monthly.csv")
RWocc <- RWocc0 %>%
    mutate(Date = mdy(Date)) %>%
    mutate(Month = as.numeric(format(Date, "%m")),
           Year = year(Date)) %>%
    # Add decimal year
    mutate(YearDec = Year + (Month - 0.5) / 12) %>%
    # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
    mutate(period = factor(period, levels = c("1", "2")),
           Site = as.factor(Site),
           SongYear = factor(SongYear, levels = paste0("Year", 1:8)),
           MonthFactor = factor(Month, levels = c(11, 12, 1:10)),
           device_type = as.factor(device_type),
           dist_200m_km = as.integer(dist_200m_km)) %>%
    # Create lagged values of PercentOccurrence, by Site
    # Group by period (not SongYear) so lags aren't reset every Nov 1, since months are
    # continuous across SongYear boundaries within a period (only period 1->2 has a real gap)
     arrange(Site, period, Date) %>%
     group_by(Site, period) %>%
     mutate(gap_days = as.numeric(Date - dplyr::lag(Date)),
            PercentOccurrence_l1 = dplyr::lag(percent_occurrence, 1, order_by = Date),
            PercentOccurrence_l2 = dplyr::lag(percent_occurrence, 2, order_by = Date),
            # invalidate lags that would bridge a real within-period data gap (>31 days)
            PercentOccurrence_l1 = if_else(!is.na(gap_days) & gap_days > 31, NA_real_, PercentOccurrence_l1),
            PercentOccurrence_l2 = if_else(dplyr::lag(gap_days, 1) > 31 | gap_days > 31, NA_real_, PercentOccurrence_l2)) %>%
     ungroup() %>%
    # Sort by Site and Date
    arrange(Site, Date) |> 
    # Create the cold pool index as a categorical factor on monthly level
    mutate(cold_pool_index = if_else(stratification >= 0.2 & bottom_temp <= 10,
                                     "Present",
                                     "Absent") |> factor(levels = c("Absent", "Present"))
    )
summary(RWocc)

# Subset without non-missing data (due to lags), if needed for modeling
RWocc_noNA <- na.omit(RWocc)
```

```{r}
# Sample sizes for different combinations
table(RWocc$Year, RWocc$Site)
table(RWocc$device_type, RWocc$period)
```


## Exploring the data on the monthly scale

```{r}
# Histogram of response variable - monthly
ggplot(RWocc, aes(x = percent_occurrence)) +
  geom_histogram(binwidth = 10, boundary = 0, fill = "lightblue", color = "black") +
  scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_y_continuous(expand = expansion(mult = c(0.03, 0.05))) +
  labs(title = "Histogram of Monthly Percent Occurrence",
       x = "Percent occurrence",
       y = "Frequency") +
  theme_minimal()
```

Count of zeros and proportion of zeros
```{r}
c(sum(RWocc$percent_occurrence == 0), 
  mean(RWocc$percent_occurrence == 0))
```

Based on the histogram and the proportion of zeros, it appears that the monthly percent occurrence data is highly skewed with a significant number of zero values. This suggests that a zero inflated distribution might be appropriate.


```{r}
# MONTHLY

# subset of data only environmental variables + percent occurrence
monthenv_vars <- RWocc[, c("Date", "SongYear", "Site", "sst", "bottom_temp", 
                           "sss", "stratification", "sstAnom",
                           "phase", "device_type", "Month", "period", "daylight_hours",
                           "cold_pool_index", "dist_200m_km", "percent_occurrence")]

# ggpairs(monthenv_vars, 
#         title = "Monthly Pairwise Correlations",
#         #mapping = aes(color = Site, alpha = 0.7),
#         upper = list(continuous = wrap("cor", size = 4.5)),
#         #lower = list(continuous = wrap("points", alpha = 0.6, size = 0.5)),
#         lower = list(combo = wrap("facethist", binwidth = 0.5))) +
#   theme_minimal()
```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWocc, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") +
  ylab("Percent occurrence")
p2 <- ggplot(RWocc, aes(x = as.factor(Month), y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") +
  labs(x = "Month", y = "Percent occurrence")
p3 <- ggplot(RWocc, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") +
  labs(x = "Period", y = "Percent occurrence")
p4 <- ggplot(RWocc, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") +
  labs(x = "Device type", y = "Percent occurrence")
p5 <- ggplot(RWocc, aes(x = as.factor(dist_200m_km), y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") +
  labs(x = "Distance (km)", y = "Percent occurrence")

# Combine with patchwork
row1 <- (p1 + p2) + plot_layout(axes = "collect_y", axis_titles = "collect_y")
row2 <- (p3 + p4 + p5) + plot_layout(axes = "collect_y", axis_titles = "collect_y")
row1 / row2 + 
  plot_annotation(title = "Monthly occurrence by different categorical variables")
```


```{r}
# Scatter plots for continuous variables
c1 <- ggplot(RWocc, aes(y = percent_occurrence, x = sst)) + geom_point() + 
  labs(x = "Sea surface temperature (°C)", y = "Percent occurrence")
c11 <- ggplot(RWocc, aes(y = percent_occurrence, x = sstAnom)) + geom_point() +
  labs(x = "Sea surface temperature anomaly (°C)", y = "Percent occurrence")
c2 <- ggplot(RWocc, aes(y = percent_occurrence, x = sss)) + geom_point() + 
  labs(x = "Sea surface salinity", y = "Percent occurrence")
c3 <- ggplot(RWocc, aes(y = percent_occurrence, x = stratification)) + geom_point() + 
  labs(x = "Stratification", y = "Percent occurrence")
c4 <- ggplot(RWocc, aes(y = percent_occurrence, x = phase)) + geom_point() + 
  labs(x = "Percent moon illumination", y = "Percent occurrence")
c5 <- ggplot(RWocc, aes(y = percent_occurrence, x = bottom_temp)) + geom_point() + 
  labs(x = "Bottom temperature (°C)", y = "Percent occurrence")

# Combine with patchwork
row1 <- (c1 + c11 + c2) + plot_layout(axes = "collect_y", axis_titles = "collect_y")
row2 <- (c3 + c4 + c5) + plot_layout(axes = "collect_y", axis_titles = "collect_y")
row1 / row2 +
  plot_annotation(title = "Monthly occurrence by different continuous variables")
 
# Interaction plot
with(RWocc,
     interaction.plot(SongYear, device_type, percent_occurrence, 
                      las = 1)
)
 
with(RWocc,
     interaction.plot(SongYear, as.factor(Month), percent_occurrence, 
                      las = 1, col = as.factor(Month))
)
```

Time series plots per site
```{r}
# Example time series plot for a specific site
ggplot(RWocc, aes(x = Date, y = percent_occurrence, 
                  group = Site)) +
  geom_line() +
  geom_point() +
  labs(title = "Monthly percent occurrence over time by site",
       x = "Date",
       y = "Percent occurrence") +
  theme_minimal() +
  facet_wrap(~ Site, scales = "free_y")
```

Time series plots per site and year
```{r}
ggplot(RWocc, aes(x = Month, y = percent_occurrence, 
                  color = SongYear,
                  group = Site)) +
  geom_line() +
  geom_point() +
  labs(title = "Monthly percent occurrence over time by site and year",
       x = "Month",
       y = "Percent occurrence") +
  theme_minimal() +
  facet_grid(SongYear ~ Site, scales = "free_y")
```



## Model from NARWC 2025

```{r}
# Modeling ---- model used in NARW consortium 2025 presentation -- monthly % occ -- year continuous
set.seed(123)
m_RWocc0 <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                    #   scs(sst, control = cs.control(cv = FALSE)) +
                      # pb(sst) +
                      # pb(sss) +
                      # pb(phase) +
                      # ga(~s(Month, bs = "cp")) +
                    #  pb(year) +
                       pb(Year) + 
                      device_type + 
                      Site
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWocc |> dplyr::select(-gap_days, -PercentOccurrence_l1, -PercentOccurrence_l2))
summary(m_RWocc0)

# year as categorical can test with anova likelihood ratio test, does make sense to use as categorical with missing 2018-2020 period rather than with smooth on continuous year.

## Model diagnostics ----
plot(m_RWocc0)
plot(m_RWocc0, ts = TRUE)
wp(m_RWocc0)


# Term plots of month and year

par(mfrow = c(1, 2))

# --- PANEL 1: Month (Custom Axis) ---
# We use terms = 1 to target the first smooth term (Month)
term.plot(m_RWocc0, terms = 1, xaxt = "n", xlab = "months")

# Define and add your custom text labels
month_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
axis(side = 1, at = 1:12, labels = month_labels)

# --- PANEL 2: Year (Default Axis) ---
# We use terms = 2 to target the second smooth term (Year)
term.plot(m_RWocc0, terms = 2, xlab = "Year")

```

## Models with Environmental Variables

#### Variables tested against percent occurrence response variable:

- **Month**, continuous with a cyclic cubic spline

- **Song Year,** Nov 1 to Oct 31

- **Sea surface temp anomalies** (sstAnom), continuous with p spline

- **Stratification** (stratification), continuous with p spline

- **Sea surface salinity** (sss), continuous with p spline

- **Percent moon illumination** (phase), continuous with p spline

- **Device type** (device_type), categorical - archival or real time

- **Site**, categorical

## Monthly Percent Occurrence Model

```{r}
# Modeling ---- monthly - year as categorical
# sstAnom vs sst chosen empirically: swapping in pb(sst) for pb(sstAnom) (all else equal)
# gives worse AIC (57.3 vs 48.2) and more autocorrelated site-period series, so sstAnom is kept.
# Site-specific seasonality via pvc(Month, by = Site) (replacing scs(Month) + Site) resolves the
# residual autocorrelation that persisted under the shared-seasonality specification.
# Zero-occurrence months cluster strongly in summer, so nu (P(exact zero) in BEINF) is modeled
# seasonally too (cyclic spline, matching the mu Month term); this removes the remaining pooled
# lag-1 autocorrelation that a season-blind nu left unexplained (AIC -66.6 vs 20.5, nu term p=0.003).
set.seed(123)
m_RWocc1 <- gamlss(percent_occurrence / 100 ~
                      pvc(Month, by = Site) +
                       pb(sstAnom) +
                       pb(stratification) +
                       pb(sss, max.df=10) + #3 for cubic relationships, 2 for parabolic
                       SongYear + 
                      device_type + 
                      PercentOccurrence_l1 + #accounts for lag
                      PercentOccurrence_l2,
                  nu.formula = ~ scs(Month, control = cs.control(cv = FALSE)), #seasonal zero-inflation
                  family = BEINF, #beta inflated
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWocc_noNA)
AIC(m_RWocc1) #-66.60372
summary(m_RWocc1)

## Model diagnostics

plot(m_RWocc1)
plot(m_RWocc1, ts = TRUE)
wp(m_RWocc1)

# Ljung-Box test per Site x period continuous series (not per SongYear, since months are
# continuous across SongYear boundaries within a period)
m_resid <- tibble(res = resid(m_RWocc1),
                Date = RWocc_noNA$Date,
                Site = RWocc_noNA$Site,
                period = RWocc_noNA$period) |>
  arrange(Site, period, Date) |> 
  group_by(Site, period) |>
  mutate(res_lag1 = dplyr::lag(res, 1, order_by = Date),
         res_lag2 = dplyr::lag(res, 2, order_by = Date)) |>
  ungroup()

m_resid |>
    group_by(Site, period) |>
    filter(n() > 4) |>
    summarise(n = n(),
              lag1_p = Box.test(res, lag = 1, type = "Ljung-Box")$p.value,
              lag2_p = Box.test(res, lag = 2, type = "Ljung-Box")$p.value,
              .groups = "drop")

# Pooling site-period series here mixes true within-series persistence with any
# regional signal shared across sites in the same calendar month, unlike the per-group tests above
cor.test(m_resid$res, m_resid$res_lag1)
cor.test(m_resid$res, m_resid$res_lag2)

# Model inference
par(mfrow = c(3, 3))
term.plot(m_RWocc1, rug = TRUE)

```

Note on getting here: without a seasonal `nu.formula`, the pooled `cor.test` above stayed
significant at lag 1 (r ~ 0.33) even though no individual Site x period series was. Tighter/
BIC-penalized mu smoothing, nonlinear `pb(PercentOccurrence_l1)`, a lag1-by-Site interaction, and
a Cochrane-Orcutt-style lagged-residual covariate all failed to remove it (some worsened per-group
results or AIC substantially). Modeling nu seasonally instead of constant removed it directly:
summer's near-total absence of whales was previously being partially absorbed into the mu residual
structure (as consecutive near-zero months), which read as autocorrelation; letting P(exact zero)
vary by Month resolves both the pooled test (lag1 p = 0.27-0.46 depending on nu smooth choice) and
all per-group tests.

Note: `stepGAIC` still further simplifies mu (dropping stratification, device_type, and
PercentOccurrence_l2) and lowers AIC to ~-75, but that simpler mu reintroduces significant
lag-1 autocorrelation for one series (T2M/period 2, p = 0.039). Since non-significant lag-1/lag-2
autocorrelation for every series is a requirement here, the fuller mu specification above is kept
as final rather than the AIC-minimal stepwise result.

## Load Biweekly Data

```{r}
# BI-WEEKLY
#setwd("~/Documents/Documents - Caroline’s MacBook Pro/Whale_Analyses/NARWCon_2025/environment_MAB_2014-2024/master data")

RWenv_biwkly <- read.csv("RWEnv_biweekly.csv")

RWenv_biwkly <- RWenv_biwkly %>%
  mutate(Date = mdy(Date)) %>%
  mutate(Month = as.numeric(format(Date, "%m")),
         Year = as.numeric(format(Date, "%Y"))) %>%
  # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
  mutate(device_type = as.factor(device_type),
         SongYear = as.factor(SongYear),
         period = as.factor(period),
         Site = as.factor(Site),
  ) %>%
     group_by(Site) %>%
     mutate(PercentOccurrence_l1 = dplyr::lag(percent_occurrence, 1, order_by = Date),
            PercentOccurrence_l2 = dplyr::lag(percent_occurrence, 2, order_by = Date)) %>%
     ungroup() %>%
    # Sort by Site and Date
    arrange(Site, Date)


# order period
RWenv_biwkly$period <- factor(RWenv_biwkly$period, levels = c("1","2"))

RWenv_biwkly$SongYear <- factor(RWenv_biwkly$SongYear, levels = c("Year1", "Year2", "Year3", "Year4", "Year5", "Year6", "Year7", "Year8"))


# Create the index as a categorical factor on monthly level - may show little
RWenv_biwkly <- RWenv_biwkly %>%
  mutate(
    cold_pool_index = if_else(
      stratification >= 0.2 & bottom_temp <= 10,
      "Present",
      "Absent"
    ),
    # Convert to a factor
    cold_pool_index = factor(cold_pool_index, levels = c("Absent", "Present"))
  )

RWenv_biwkly <- na.omit(RWenv_biwkly)

RWenv_biwkly$PercentOccurrence_l1 <- as.numeric(RWenv_biwkly$PercentOccurrence_l1)
RWenv_biwkly$PercentOccurrence_l2 <- as.numeric(RWenv_biwkly$PercentOccurrence_l2)
```

## Exploring the data on the biweekly scale

```{r}
# BI WEEKLY 

#subset of data only environmental variables + percent occurrence
biwkenv_vars <- RWenv_biwkly[, c("Date", "Site", "sst", "bottom_temp","sss","stratification", 
                        "sstAnom", "phase", "device_type", "Month","period", "daylight_hours",
                           "cold_pool_index","percent_occurrence")]

ggpairs(biwkenv_vars, 
        title = "Biweekly Pairwise Correlations",
         #mapping = aes(color = Site, alpha = 0.7),
        upper = list(continuous = wrap("cor", size = 4.5)),
       lower = list(combo = wrap("facethist", binwidth = 0.5))) +
  theme_minimal()

```

```{r}
# Histogram of response variable
hist(RWenv_biwkly$percent_occurrence)
```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWenv_biwkly, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWenv_biwkly, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWenv_biwkly, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p4 <- ggplot(RWenv_biwkly, aes(x = SongYear, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2 ) / ( p3 | p4) +
  plot_annotation(title = "Biweekly occurrence by different categorical variables")



```

## Biweekly Percent Occurrence Model

Introduction of lagged biweekly variables 'PercentOccurrence_l1' for previous 2-week period and 'PercentOccurrence_l2' for subsequent 2-week period and corARMA correlation structure in model.

```{r}
# Modeling ---- biweekly - year as categorical

set.seed(123)
m_RWoccbiwk <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                    #   scs(sst, control = cs.control(cv = FALSE)) +
                       pb(sstAnom) +
                        pb(stratification) +
                       pb(sss,max.df =9) +
                       pb(phase, max.df =8, control = pb.control(inter = 5)) +
                      # ga(~s(Month, bs = "cp")) +
                     # pb(Year) +
                       SongYear +  
                      device_type + 
                    cold_pool_index +
                      Site +
                     PercentOccurrence_l1 + PercentOccurrence_l2,
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                 correlation = corARMA(p = 1),
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWenv_biwkly)
summary(m_RWoccbiwk)

# Select model based on AIC
m_RWoccbiwk <- gamlss::stepGAIC(m_RWoccbiwk)

## Model diagnostics ----
plot(m_RWoccbiwk)
plot(m_RWoccbiwk, ts = TRUE)
wp(m_RWoccbiwk, ylim.all=3)

term.plot(m_RWoccbiwk, rug = TRUE)
```

## Load Weekly Data

```{r}
#WEEKLY
#setwd("~/Documents/Documents - Caroline’s MacBook Pro/Whale_Analyses/NARWCon_2025/environment_MAB_2014-2024/master data")

RWenv_wkly <- read.csv("RWEnv_weekly.csv")

RWenv_wkly <- RWenv_wkly %>%
  mutate(Date = mdy(Date)) %>%
  mutate(Month = as.numeric(format(Date, "%m")),
         Year = as.numeric(format(Date, "%Y"))) %>%
  # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
  mutate(device_type = as.factor(device_type),
         SongYear = as.factor(SongYear),
         period = as.factor(period),
         Site = as.factor(Site),
        ) %>%
     group_by(Site) %>%
     mutate(PercentOccurrence_l1 = dplyr::lag(percent_occurrence, 1, order_by = Date),
            PercentOccurrence_l2 = dplyr::lag(percent_occurrence, 2, order_by = Date)) %>%
     ungroup() %>%
  # Sort by Site and Date
  arrange(Site, Date)



# order period
RWenv_wkly$period <- factor(RWenv_wkly$period, levels = c("1","2"))

# Change year to categorical
RWenv_wkly$SongYear <- factor(RWenv_wkly$SongYear, levels = c("Year1", "Year2", "Year3", "Year4", "Year5", "Year6", "Year7", "Year8"))


# Create the index as a categorical factor on monthly level - may show little
RWenv_wkly <- RWenv_wkly %>%
  mutate(
    cold_pool_index = if_else(
      stratification >= 0.2 & bottom_temp <= 10,
      "Present",
      "Absent"
    ),
    # Convert to a factor
    cold_pool_index = factor(cold_pool_index, levels = c("Absent", "Present")))

RWenv_wkly <- na.omit(RWenv_wkly)

```

## Exploring the data on the weekly scale

```{r}
# WEEKLY 

#subset of data only environmental variables + percent occurrence
wkenv_vars <- RWenv_wkly[, c("Date", "Site", "sst", "bottom_temp", "sss","stratification",
                            "phase", "device_type", "Month","period" , "daylight_hours",
                           "cold_pool_index", "sstAnom", "percent_occurrence")]

ggpairs(wkenv_vars, 
        title = "Weekly Pairwise Correlations",
         #mapping = aes(color = Site, alpha = 0.7),
        upper = list(continuous = wrap("cor", size = 4.5)),
        lower = list(combo = wrap("facethist", binwidth = 0.5)))  +
  theme_minimal()

```

```{r}
# Histogram of response variable
hist(RWenv_wkly$percent_occurrence)

```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWenv_wkly, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWenv_wkly, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWenv_wkly, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2 | p3) +
  plot_annotation(title = "Weekly occurrence by different categorical variables")

```

## Weekly Percent Occurrence Model

Lagged biweekly variables 'PercentOccurrence_l1' for previous week and 'PercentOccurrence_l2' for subsequent week and corARMA correlation structure in model.

```{r}
# Modeling --- weekly - year as categorical

set.seed(123)
m_RWoccwk <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                      # scs(sst, control = cs.control(cv = FALSE)) +
                       pb(sstAnom) +
                        pb(stratification) +
                       pb(sss,max.df =9, inter=10) + 
                       pb(phase,max.df =9, inter=10) +
                      # ga(~s(Month, bs = "cp")) +
                     # pb(Year) +
                       SongYear +  # seems like year as continuous is better
                      device_type +
                      Site +
                     cold_pool_index +
                     PercentOccurrence_l1 + PercentOccurrence_l2,
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                  correlation = corARMA(p = 1),
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWenv_wkly)
summary(m_RWoccwk)

# Select model based on AIC
m_RWoccwk <- gamlss::stepGAIC(m_RWoccwk)
```

```{r}
## Model diagnostics ----
plot(m_RWoccwk)
plot(m_RWoccwk, ts = TRUE)
wp(m_RWoccwk, ylim.all=3)

term.plot(m_RWoccwk, rug = TRUE)
```
