---
title: "NARW_GAM_Models"
format: html
self-contained: true
editor: visual
fig-width: 12
fig-height: 12
out-width: "100%"
toc: true
---

# NARW Occurrence GAM Model Testing

## Packages

```{r}
library(dplyr)
library(lubridate)
# library(scales)

# Visualization
library(ggplot2)
theme_set(theme_light())
library(patchwork)
library(GGally)

# Modeling
library(gamlss)
# library(gamlss.tr)
library(gamlss.add)
 library(mgcv)
```

## Load Monthly Data

```{r}
# Load data ----
#setwd("~/Documents/Documents - Caroline’s MacBook Pro/Whale_Analyses/NARWCon_2025/environment_MAB_2014-2024/master data")

RWocc0 <- read.csv("RWEnv_monthly.csv")
RWocc <- RWocc0 %>%
    mutate(Date = mdy(Date)) %>%
    mutate(Month = as.numeric(format(Date, "%m"))) %>%
    # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
    mutate(MonthCat = as.factor(month),
           period = as.factor(period),
           Site = as.factor(Site),
           device_type = as.factor(device_type),
           dist_200m_km = as.integer(dist_200m_km)) %>%
    # Create lagged values of PercentOccurrence, by Site
    # group_by(Site) %>%
    # mutate(PercentOccurrence_l1 = dplyr::lag(PercentOccurrence, 1, order_by = Date),
    #        PercentOccurrence_l2 = dplyr::lag(PercentOccurrence, 2, order_by = Date)) %>%
    # ungroup() %>%
    # Sort by Site and Date
    arrange(Site, Date)

#order period
RWocc$period <- factor(RWocc$period, levels = c("1","2"))

# Change year to categorical
RWocc$YearCat <- as.factor(RWocc$year)
# Showing year is in chronological order - 2014 only includes 11/3 to 12/31 of that year
RWocc$YearCat <- factor(RWocc$year, levels = c("2014", "2015", "2016", "2017", "2021", "2022", "2023", "2024"))


# Create the index as a categorical factor on monthly level - may show little
RWocc <- RWocc %>%
  mutate(
    cold_pool_index = if_else(
      stratification >= 0.2 & bottom_temp <= 10,
      "Present",
      "Absent"
    ),
    # Convert to a factor
    cold_pool_index = factor(cold_pool_index, levels = c("Absent", "Present"))
  )
```

```{r}
# Sample sizes for different combinations
table(RWocc$Year, RWocc$Site)
table(RWocc$device_type, RWocc$period)
```

## Exploring the data on the monthly scale

```{r}
# Histogram of response variable - monthly
hist(RWocc$percent_occurrence)

#zero inflated distribution
```

```{r}
# MONTHLY

#subset of data only environmental variables + percent occurrence
monthenv_vars <- RWocc[, c("Date", "Site", "sst", "bottom_temp", "sss", "stratification",
                           "phase", "device_type", "Month","period" , "daylight_hours",
                          "cold_pool_index", "dist_200m_km", "percent_occurrence")]

ggpairs(monthenv_vars, 
        title = "Monthly Pairwise Correlations",
         #mapping = aes(color = Site, alpha = 0.7),
        upper = list(continuous = wrap("cor", size = 4.5)),
        #lower = list(continuous = wrap("points", alpha = 0.6, size = 0.5)),
        lower = list(combo = wrap("facethist", binwidth = 0.5))) +
  theme_minimal()

```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWocc, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWocc, aes(x = as.factor(month), y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWocc, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p4 <- ggplot(RWocc, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p5 <- ggplot(RWocc, aes(x = as.factor(dist_200m_km), y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2) / (p3 | p4 | p5) +
  plot_annotation(title = "Monthly occurrence by different categorical variables")


# Scatter plots for continuous variables
c1 <- ggplot(RWocc, aes(y=percent_occurrence, x=sst)) + geom_point()
c2 <-  ggplot(RWocc, aes(y=percent_occurrence, x=sss)) + geom_point()
c3 <-  ggplot(RWocc, aes(y=percent_occurrence, x=stratification)) + geom_point()
c4 <- ggplot(RWocc, aes(y=percent_occurrence, x=phase)) + geom_point()
c5 <- ggplot(RWocc, aes(y=percent_occurrence, x=bottom_temp)) + geom_point()
# Combine with patchwork
(c1 | c2) / (c3 | c4 | c5) +
  plot_annotation(title = "Monthly occurrence by different continuous variables")

 
 # Interaction plot
 with(RWocc,
     interaction.plot(YearCat, device_type, percent_occurrence)
)
 with(RWocc,
     interaction.plot(YearCat, MonthCat, percent_occurrence, col = MonthCat)
)


```

## Model from NARWC 2025

```{r}
# Modeling ---- model used in NARW consortium 2025 presentation -- monthly % occ -- year continuous
set.seed(123)
m_RWocc0 <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                    #   scs(sst, control = cs.control(cv = FALSE)) +
                      # pb(sst) +
                      # pb(sss) +
                      # pb(phase) +
                      # ga(~s(Month, bs = "cp")) +
                    #  pb(year) +
                       YearCat + # seems like year as continuous is better
                      device_type + 
                      Site
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWocc)
summary(m_RWocc0)

# year as categorical can test with anova likelihood ratio test, does make sense to use as categorical with missing 2018-2020 period rather than with smooth on continuous year.

## Model diagnostics ----
plot(m_RWocc0)
plot(m_RWocc0, ts = TRUE)
wp(m_RWocc0)


# Term plots of month and year

par(mfrow = c(1, 2))

# --- PANEL 1: Month (Custom Axis) ---
# We use terms = 1 to target the first smooth term (Month)
term.plot(m_RWocc0, terms = 1, xaxt = "n", xlab = "months")

# Define and add your custom text labels
month_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
axis(side = 1, at = 1:12, labels = month_labels)

# --- PANEL 2: Year (Default Axis) ---
# We use terms = 2 to target the second smooth term (Year)
term.plot(m_RWocc0, terms = 2, xlab = "Year")

```

## Models with Environmental Variables

#### Variables tested against percent occurrence response variable:

-   **Month**, continuous with a cyclic cubic spline

-   **Year** (YearCat), categorical ordered factor (chronological)

    -   Chose not to run year as continuous due to the 2018-2020 gap period, feel it makes more sense to use as categorical. Previously we re-ordered the data to start at 2015 as the year to compare with, here I have left it in chronological order.

-   **Sea surface temp** (sst), continuous with p spline

-   **Sea surface salinity** (sss), continuous with p spline

-   **Percent moon illumination** (phase), continuous with p spline

-   **Device type** (device_type), categorical - archival or real time

-   **Site**, categorical

## Monthly Percent Occurrence Model

```{r}
# Modeling ---- monthly - year as categorical
set.seed(123)
m_RWocc1 <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                    #   scs(sst, control = cs.control(cv = FALSE)) +
                       pb(sst) +
                       pb(sss, max.df=10) + #3 for cubic relationships, 2 for parabolic
                       pb(phase) +
                      # ga(~s(Month, bs = "cp")) +
                    #  pb(year) +
                       YearCat + # seems like year as continuous is better
                      device_type + 
                      Site 
                   #   period
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWocc)
summary(m_RWocc1)

# Select model based on AIC
m_RWocc1 <- gamlss::stepGAIC(m_RWocc1)
    # drops sst 

## Model diagnostics ----
plot(m_RWocc1)
plot(m_RWocc1, ts = TRUE)
wp(m_RWocc1)


term.plot(m_RWocc1, rug = TRUE)

```

## Load Biweekly Data

```{r}
# BI-WEEKLY
#setwd("~/Documents/Documents - Caroline’s MacBook Pro/Whale_Analyses/NARWCon_2025/environment_MAB_2014-2024/master data")

RWenv_biwkly <- read.csv("RWEnv_biweekly.csv")

RWenv_biwkly <- RWenv_biwkly %>%
  mutate(Date = mdy(Date)) %>%
  mutate(Month = as.numeric(format(Date, "%m")),
         Year = as.numeric(format(Date, "%Y"))) %>%
  # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
  mutate(device_type = as.factor(device_type),
         YearCat = as.factor(Year),
         period = as.factor(period),
         Site = as.factor(Site),
  ) %>%
     group_by(Site) %>%
     mutate(PercentOccurrence_l1 = dplyr::lag(percent_occurrence, 1, order_by = Date),
            PercentOccurrence_l2 = dplyr::lag(percent_occurrence, 2, order_by = Date)) %>%
     ungroup() %>%
    # Sort by Site and Date
    arrange(Site, Date)


# order period
RWenv_biwkly$period <- factor(RWenv_biwkly$period, levels = c("1","2"))

# Change year to categorical
RWenv_biwkly$YearCat <- as.factor(RWenv_biwkly$Year)
RWenv_biwkly$YearCat <- factor(RWenv_biwkly$Year, levels = c("2014", "2015", "2016", "2017", "2021", "2022", "2023", "2024"))


# Create the index as a categorical factor on monthly level - may show little
RWenv_biwkly <- RWenv_biwkly %>%
  mutate(
    cold_pool_index = if_else(
      stratification >= 0.2 & bottom_temp <= 10,
      "Present",
      "Absent"
    ),
    # Convert to a factor
    cold_pool_index = factor(cold_pool_index, levels = c("Absent", "Present"))
  )

RWenv_biwkly <- na.omit(RWenv_biwkly)

RWenv_biwkly$PercentOccurrence_l1 <- as.numeric(RWenv_biwkly$PercentOccurrence_l1)
RWenv_biwkly$PercentOccurrence_l2 <- as.numeric(RWenv_biwkly$PercentOccurrence_l2)
```

## Exploring the data on the biweekly scale

```{r}
# BI WEEKLY 

#subset of data only environmental variables + percent occurrence
biwkenv_vars <- RWenv_biwkly[, c("Date", "Site", "sst", "bottom_temp","sss","stratification", 
                                 "phase", "device_type", "Month","period", "daylight_hours",
                           "cold_pool_index","percent_occurrence")]

ggpairs(biwkenv_vars, 
        title = "Biweekly Pairwise Correlations",
         #mapping = aes(color = Site, alpha = 0.7),
        upper = list(continuous = wrap("cor", size = 4.5)),
       lower = list(combo = wrap("facethist", binwidth = 0.5))) +
  theme_minimal()

```

```{r}
# Histogram of response variable
hist(RWenv_biwkly$percent_occurrence)
```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWenv_biwkly, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWenv_biwkly, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWenv_biwkly, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2 | p3) +
  plot_annotation(title = "Biweekly occurrence by different categorical variables")


```

## Biweekly Percent Occurrence Model

Introduction of lagged biweekly variables 'PercentOccurrence_l1' for previous 2-week period and 'PercentOccurrence_l2' for subsequent 2-week period and corARMA correlation structure in model.

```{r}
# Modeling ---- biweekly - year as categorical

set.seed(123)
m_RWoccbiwk <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                    #   scs(sst, control = cs.control(cv = FALSE)) +
                       pb(sst) +
                       pb(sss,max.df =9) +
                       pb(phase, max.df =9, control = pb.control(inter = 6)) +
                      # ga(~s(Month, bs = "cp")) +
                     # pb(Year) +
                       YearCat +  # seems like year as continuous is better
                      device_type + 
                      Site
                    + PercentOccurrence_l1 + PercentOccurrence_l2,
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                 correlation = corARMA(p = 1),
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWenv_biwkly)
summary(m_RWoccbiwk)

# Select model based on AIC
m_RWoccbiwk <- gamlss::stepGAIC(m_RWoccbiwk)

## Model diagnostics ----
plot(m_RWoccbiwk)
plot(m_RWoccbiwk, ts = TRUE)
wp(m_RWoccbiwk, ylim.all=3)

term.plot(m_RWoccbiwk, rug = TRUE)
```

## Load Weekly Data

```{r}
#WEEKLY
#setwd("~/Documents/Documents - Caroline’s MacBook Pro/Whale_Analyses/NARWCon_2025/environment_MAB_2014-2024/master data")

RWenv_wkly <- read.csv("RWEnv_weekly.csv")

RWenv_wkly <- RWenv_wkly %>%
  mutate(Date = mdy(Date)) %>%
  mutate(Month = as.numeric(format(Date, "%m")),
         Year = as.numeric(format(Date, "%Y"))) %>%
  # Change Site, Month, and Period (1 = 2014-17, 2 = 2021-2024) to categorical
  mutate(device_type = as.factor(device_type),
         YearCat = as.factor(Year),
         period = as.factor(period),
         Site = as.factor(Site),
        ) %>%
     group_by(Site) %>%
     mutate(PercentOccurrence_l1 = dplyr::lag(percent_occurrence, 1, order_by = Date),
            PercentOccurrence_l2 = dplyr::lag(percent_occurrence, 2, order_by = Date)) %>%
     ungroup() %>%
  # Sort by Site and Date
  arrange(Site, Date)



# order period
RWenv_wkly$period <- factor(RWenv_wkly$period, levels = c("1","2"))

# Change year to categorical
RWenv_wkly$YearCat <- as.factor(RWenv_wkly$Year)
RWenv_wkly$YearCat <- factor(RWenv_wkly$Year, levels = c("2014", "2015", "2016", "2017", "2021", "2022", "2023", "2024"))


# Create the index as a categorical factor on monthly level - may show little
RWenv_wkly <- RWenv_wkly %>%
  mutate(
    cold_pool_index = if_else(
      stratification >= 0.2 & bottom_temp <= 10,
      "Present",
      "Absent"
    ),
    # Convert to a factor
    cold_pool_index = factor(cold_pool_index, levels = c("Absent", "Present")))

RWenv_wkly <- na.omit(RWenv_wkly)

```

## Exploring the data on the weekly scale

```{r}
# WEEKLY 

#subset of data only environmental variables + percent occurrence
wkenv_vars <- RWenv_wkly[, c("Date", "Site", "sst", "bottom_temp", "sss","stratification",
                            "phase", "device_type", "Month","period" , "daylight_hours",
                           "cold_pool_index", "percent_occurrence")]

ggpairs(wkenv_vars, 
        title = "Weekly Pairwise Correlations",
         #mapping = aes(color = Site, alpha = 0.7),
        upper = list(continuous = wrap("cor", size = 4.5)),
        lower = list(combo = wrap("facethist", binwidth = 0.5)))  +
  theme_minimal()

```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWenv_wkly, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWenv_wkly, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWenv_wkly, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2 | p3) +
  plot_annotation(title = "Weekly occurrence by different categorical variables")

```

```{r}
# Histogram of response variable
hist(RWenv_wkly$percent_occurrence)

```

```{r}
# Boxplots of Occurrence by different categorical variables
p1 <- ggplot(RWenv_wkly, aes(x = Site, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue") 
p2 <- ggplot(RWenv_wkly, aes(x = period, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
p3 <- ggplot(RWenv_wkly, aes(x = device_type, y = percent_occurrence)) +
  geom_boxplot(fill = "lightblue")
# Combine with patchwork
(p1 | p2 | p3) +
  plot_annotation(title = "Weekly occurrence by different categorical variables")

```

## Weekly Percent Occurrence Model

Lagged biweekly variables 'PercentOccurrence_l1' for previous week and 'PercentOccurrence_l2' for subsequent week and corARMA correlation structure in model.

```{r}
# Modeling --- weekly - year as categorical

set.seed(123)
m_RWoccwk <- gamlss(percent_occurrence / 100 ~
                    #  ga(~ti(year, Month)) +
                      scs(Month, control = cs.control(cv = FALSE)) +
                      # scs(sst, control = cs.control(cv = FALSE)) +
                       pb(sst) +
                       pb(sss,max.df =9, inter=10) + 
                       pb(phase,max.df =9, inter=10) +
                      # ga(~s(Month, bs = "cp")) +
                     # pb(Year) +
                       YearCat +  # seems like year as continuous is better
                      device_type +
                      Site 
                    + PercentOccurrence_l1 + PercentOccurrence_l2,
                  # + pvc(Month, by = Site)
                  # ,sigma.formula = ~pb(Month) #+ pb(Year)
                  ,family = BEINF, #beta inflated
                  correlation = corARMA(p = 1),
                  control = gamlss.control(c.crit = 0.01, n.cyc = 100),
                  data = RWenv_wkly)
summary(m_RWoccwk)

# Select model based on AIC
m_RWoccwk <- gamlss::stepGAIC(m_RWoccwk)
```

```{r}
## Model diagnostics ----
plot(m_RWoccwk)
plot(m_RWoccwk, ts = TRUE)
wp(m_RWoccwk, ylim.all=3)

term.plot(m_RWoccwk, rug = TRUE)
```
